Toronto Beach Data

### 2024-09-27 paper/paper.qmd
First Part of Data Completed

---
title: "Toronto Beach Data"
subtitle: "Data Analysis of the Relationship Between Date of Observation, Air Temperatures, Water Temperatures, Rainfall Presence, and waterfowl activity"
author: "Yanzun Jiang"
date: today
date-format: long
thanks: "Code and data supporting this analysis is available at: <https://github.com/Stary54264/Toronto-Beach-Data>"
abstract: "This report analyzes environmental data collected in the summer of 2023 regarding Toronto's beaches, focusing on variables such as air temperature, water temperature, rainfall, and waterfowl observations. The findings reveal that #TODO increased water temperatures correlate with higher waterfowl activity, while rainy conditions tend to reduce sightings.# These insights emphasize the importance of understanding how weather patterns affect wildlife behavior and beach usability, suggesting that public awareness and management strategies should be adapted accordingly. Future research should explore the impact of these environmental factors on beachgoer experiences and the broader ecological implications."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

Here is a preview of the dataset used in this analysis @tbl-data, which spans from 2008 to present. The dataset provides a comprehensive overview of key environmental variables observed at Toronto's beaches, including the date of observation, air and water temperatures, rainfall presence, and waterfowl activity. These variables are crucial for understanding long-term trends in environmental conditions and their impact on beach safety, water quality, and wildlife behavior. This preview serves as an introduction to the dataset's structure and quality, setting the stage for in-depth analysis and trend exploration over time.

@@

### Air Temperature

This histogram (@fig-air-temp) shows the distribution of air temperatures of beaches in Toronto.

```{r}
#| label: fig-air-temp
#| fig-cap: Air Temperature at Toronto Beaches from
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = air_temp)) +
  geom_histogram(bins = 30, col = "darkblue", fill = "lightblue") +
  theme_minimal() +
  labs(x = "Air Temperature", y = "# of Occurences")
```

We can clearly see the data is unimodal and is not skewed.

### Rain

This bar graph (@fig-rain) shows the distribution of whether it rains of not at beaches in Toronto.

```{r}
#| label: fig-rain
#| fig-cap: Bar Graph of Whether It Rains or Not
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = rain)) +
  geom_bar(col = "darkblue", fill = "lightblue") +
  theme_minimal() +
  labs(x = "Whether It Rains or Not", y = "# of Occurences")
```

We can clearly see that there are less raining days.

### Water Temperature

This histogram (@fig-water-temp) shows the distribution of water temperatures of beaches in Toronto.

```{r}
#| label: fig-water-temp
#| fig-cap: Histogram of Water Temperature
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = water_temp)) +
  geom_histogram(bins = 30, col = "darkblue", fill = "lightblue") +
  theme_minimal() +
  labs(x = "Water Temperature", y = "# of Occurences")
```

We can clearly see the data is unimodal, which is the same as air temperatures, but it is left-skewed.

### Observations of Water Fowls

This histogram (@fig-water-fowl) shows the distribution of observations of water fowls of beaches in Toronto.

```{r}
#| label: fig-water-fowl
#| fig-cap: Histogram of Observations of Water Fowls
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = water_fowl)) +
  geom_histogram(bins = 30, col = "darkblue", fill = "lightblue") +
  theme_minimal() +
  labs(x = "Observations of Water Fowls", y = "# of Occurences")
```

We can clearly see the data is unimodal but highly right-skewed.

@@

Our results are summarized in.

@@

library(rstanarm)
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

@@

The data cleaning process involved renaming variables, filtering out useless columns from the raw dataset, filtering out observations with NAs, and removing the outliers using z-scores.

>>>>

---
title: "Toronto Beach Data"
subtitle: "Data Analysis of the Relationship Between Date of Observation, Air Temperatures, Rainfall Presence, Water Temperatures, and Waterfowl Activities"
author: "Yanzun Jiang"
date: today
date-format: long
thanks: "Code and data supporting this analysis is available at: <https://github.com/Stary54264/Toronto-Beach-Data>"
abstract: "This report analyzes environmental data collected in the summer of 2023 regarding Toronto's beaches, focusing on variables such as air temperature, water temperature, rainfall, and waterfowl observations. The findings reveal that #TODO increased water temperatures correlate with higher waterfowl activity, while rainy conditions tend to reduce sightings.# These insights emphasize the importance of understanding how weather patterns affect wildlife behavior and beach usability, suggesting that public awareness and management strategies should be adapted accordingly. Future research should explore the impact of these environmental factors on beachgoer experiences and the broader ecological implications."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

The preview of the dataset (@tbl-data) spans from 2008 to present. The dataset provides a comprehensive overview of key environmental variables observed at Toronto's beaches, including the date of observation, air and water temperatures, rainfall presence, and waterfowl activity. These variables are crucial for understanding long-term trends in environmental conditions and their impact on beach safety, water quality, and wildlife behavior. This preview serves as an introduction to the dataset's structure and quality, setting the stage for in-depth analysis and trend exploration over time.

@@

### Distribution of the Variables

These graphs (@fig-airtemp, @fig-rain, @fig-watertemp, and @fig-waterfowl) visualizes the distribution of air temperatures, rainfall observations, water temperature, and waterfowl observations across the dataset from 2008 to the present. These graphs helps identify patterns, outliers, and the overall spread of the data. By examining these distributions, we can gain insights into how they behave over time and how it may influence other variables, such as water quality or wildlife observations.

```{r}
#| label: fig-airtemp
#| fig-cap: Distribution of Air Temperature at Toronto Beaches
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = air_temp)) +
  geom_histogram(bins = 30, col = "darkblue", fill = "lightblue") +
  theme_minimal() +
  labs(x = "Air Temperature (°C)", y = "Frequency")
```

```{r}
#| label: fig-rain
#| fig-cap: Rainfall Observations at Toronto Beaches
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = rain)) +
  geom_bar(col = "darkblue", fill = "lightblue") +
  theme_minimal() +
  labs(x = "Rainfall Observed", y = "Frequency")
```

```{r}
#| label: fig-watertemp
#| fig-cap: Distribution of Water Temperature at Toronto Beaches
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = water_temp)) +
  geom_histogram(bins = 30, col = "darkblue", fill = "lightblue") +
  theme_minimal() +
  labs(x = "Water Temperature (°C)", y = "Frequency")
```

```{r}
#| label: fig-waterfowl
#| fig-cap: Distribution of Waterfowl Observations at Toronto Beaches
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = water_fowl)) +
  geom_histogram(bins = 30, col = "darkblue", fill = "lightblue") +
  theme_minimal() +
  labs(x = "Number of Waterfowl Observed", y = "Frequency")
```

## Relationships Between the Variables

### Changes of Beaches with Time

These graphs (@fig-time-airtemp-year, @fig-time-rain, @fig-time-watertemp-year, and @fig-time-waterfowl-year) visualize the relationship between each environmental variable and time, spanning from 2008 to the present. By plotting air temperature, rainfall observations, water temperature, and waterfowl observations against time, we aim to identify trends, seasonal patterns, and potential long-term changes in these variables at Toronto's beaches. Each scatter plot provides insights into how these environmental factors have fluctuated over the years, revealing trends in beach weather conditions and wildlife presence. These time-based visualizations will aid in understanding any periodic or gradual changes and their potential impact on beach safety and ecological health.

```{r}
#| label: fig-time-airtemp-year
#| fig-cap: Air Temperature Over Time at Toronto Beaches
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = date, y = air_temp)) +
  geom_point(col = "lightblue") +
  theme_minimal() +
  labs(x = "Date", y = "Air Temperature (°C)")
```

```{r}
#| label: fig-time-rain
#| fig-cap: Rainfall Observations Over Time at Toronto Beaches
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = date, y = rain)) +
  geom_boxplot(col = "darkblue", fill = "lightblue") +
  theme_minimal() +
  labs(x = "Date", y = "Rainfall")
```

```{r}
#| label: fig-time-watertemp-year
#| fig-cap: Water Temperature Over Time at Toronto Beaches
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = date, y = water_temp)) +
  geom_point(col = "lightblue") +
  theme_minimal() +
  labs(x = "Date", y = "Water Temperature (°C)")
```

```{r}
#| label: fig-time-waterfowl-year
#| fig-cap: Waterfowl Observations Over Time at Toronto Beaches
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = date, y = water_fowl)) +
  geom_point(col = "lightblue") +
  theme_minimal() +
  labs(x = "Date", y = "Number of Waterfowl Observed")
```

@@

From @fig-airtemp, we can clearly see the data is unimodal and is not skewed.

From @fig-rain, we can clearly see that there are less raining days than non-raining days.

From @fig-watertemp, we can clearly see the data is unimodal, which is the same as air temperatures, but it is left-skewed.

From @fig-waterfowl, we can clearly see the data is unimodal but highly right-skewed.

@@

@@

The data cleaning process involved renaming variables, tidying the dates, filtering out useless columns from the raw dataset, filtering out observations with NAs, and removing the outliers using z-scores.
